Discussion Draft

### 2020-04-06 doc/Discussion.rmd
better lines for interference part

The **COM** model predicts that interference and the recombination landscape arises from known oscillatory movements during prophase, it lacks evolutionary based predictions.

While the **SACE** modifier model does not explicitly model evolution of interference strength we note that a logical outcome of the main prediction of maintaining larger chromosome blocks in males, would be a landscape with stronger interference strength.

 We propose the **spindle based selection model** would support the evolution of interference strength in the positive direction via modulation of the amount of sister cohesion connecting homologs (figure).

<**other models**>

Models from Goldstein et al (review in [@ottoPaysuer2019] suggest that if this pattern is widespread interference evolves whenever increased recombination rates evolve. Perhaps a distinguishing feature of models which come to this finding is that the number of crossovers is kept constant. The space across multiple loci (veller?) or between multiple crossovers increases in a positive manner with genome wide recombination rates. Given that the empirical range of crossovers per chromosome is quite small (1-3 [@ottoPaysuer2019, @stapley_variation_2017] and  the obligate crossover rule, the assumption of constraining the number of crossovers per chromosome fits well with empirical data.

<Biological exceptions to interference:  experimentally increasing crossover number through mutants or fusion chromosomes  (Celegans, (plants, fungi with negative interference.>

#Future Steps

```{r future.step.plan, echo=FALSE}
#still preliminary
```

- sex-specific and sex average biases. Some cytology models (Celegans) mostly focus on oogenesis  and genetic maps are often from hermaphrodites. (lack of data from one sex). Recognize the value of Chromosome level measures and incorporating chromosome behavior into models.

Broader communication across sub-fields of cell physiology and biophysics for understanding recombination rate evolution. Utilize mitosis systems given the conservation of the molecular pathway. Think of meiosis as a program.

Dark matter of the genome. We can't really exclude that some of the recombination rate variation is due to structural variation (more of less genome). (since everything is mapped to B6). (also maybe there are crossovers in the centromeres and telomeres -- which might escape detected in linkage maps)

what are the rates of sister-sister crossovers?

How does elevated rates of precocious sister centromere separation effect oogenesis?

Are selfish elements affecting the evolution of meiosis -- and thus recombination rates (from a functional standpoint).

Keep generating those testable hypotheses (cite).

- chromatin interference

>>>>

The **COM** model predicts that interference and the recombination landscape arises from known oscillatory movements during prophase, however it lacks evolutionary based predictions. While the **SACE** modifier model does not explicitly model evolution of interference strength we note that a logical outcome of the main prediction of maintaining larger chromosome blocks in males, would be a landscape with stronger interference strength.

 We propose that the **spindle based selection model** support the evolution of stronger interference strength via selection onthe amount of sister cohesion connecting homologs (figure).  -- and this can evolve with increased crossover number

<we don't know the chromosome identities, but based on the propotions of chromosome classes this is the logical conclusion.

The evolution of gwRR in our study is driven by a transition of recombination landscapes in a propotion of chromosomes, from single crossovers to two crossovers --

All else being equal, axis length) -- interference strength would be expected to be equal or even weaker

Our model suggests that an outcome of this further spacing of two foci results in an increased area of sister cohesion connecting homologs and could change the chromsome structure of these bivalents when they are aligned and bi-oriented. This model gives rise to the possibility that the chromosome structure at metaphase is the target of selection and changing the recombination landscape is an outcome  pleiotropic. This interpertation has potential to fit with hypotheses mentioned above for explaining other results regarding the evolution of the recombination landscape.

We have focused on the model involving direct selection on the recombination pathway, that is the impact on the meotic process on an indiviual, but we make note on models centering on indirect selection or fitness outcomes of offspring which could also explain our results. Models from Goldstein et al, reviewed in [@ottoPaysuer2019] and Veller 2017?, predict that interference strength evolves whenever increased recombination rates evolve. The simulations under this model indicate that the space across multiple loci or between multiple crossovers increases in a positive manner with the genome wide recombination rate. A distinguishing feature of their model is that the number of crossovers per chromosome (genetic unit) is kept constant. The range of crossovers per chromosome is quite small, 1-3, for the majority of chromosomes [@ottoPaysuer2019, @stapley_variation_2017]. Constraining the number of crossovers per chromosome in models may fit empiracal data better than those where recombiantion rate across an abstracted genetic space is unconstrained.

```{r future.step.plan, echo=FALSE, include=FALSE}
#still preliminary
#sex-specific and sex average biases. Some cytology models (Celegans) mostly focus on oogenesis  and genetic maps are often from hermaphrodites. (lack of data from one sex). Recognize the value of Chromosome level measures and incorporating chromosome behavior into models.
#Broader communication across sub-fields of cell physiology and biophysics for understanding recombination rate evolution. Utilize mitosis systems given the conservation of the molecular pathway. Think of meiosis as a program.

#Dark matter of the genome. We can't really exclude that some of the recombination rate variation is due to structural variation (more of less genome). (since everything is mapped to B6). (also maybe there are crossovers in the centromeres and telomeres -- which might escape detected in linkage maps)

#what are the rates of sister-sister crossovers?
#How does elevated rates of precocious sister centromere separation effect oogenesis?

#Are selfish elements affecting the evolution of meiosis -- and thus recombination rates (from a functional standpoint).

#Keep generating those testable hypotheses (cite).  
#- chromatin interference
```
